---
title: "Yamamoto Study"
author: "Akihiro Shiroshita"
date: "`r Sys.time()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	fig.height = 4,
	fig.pos = "t",
	message = FALSE,
	warning = FALSE,
	dpi = 350,
	out.extra = ""
)
packages = c("data.table",
             "lava",
             "lme4",
             "nlme",
             "readxl",
             "irr",
             "tidyverse",
             "psych",
             "lubridate",
             "ggplot2",
             "RColorBrewer",
             "ggplotgui",
             "boot",
             "knitr")
package.check <- lapply(packages, FUN = function(x){
  if (!require(x, character.only = TRUE)){
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

```

```{r}
main_gai <- read_excel("input/data.xls", sheet = "main_gai")
main_vtm <- read_excel("input/data.xls", sheet = "main_vtm")
sub_over_gai <- read_excel("input/data.xls", sheet = "sub_over_gai")
sub_over_vtm <- read_excel("input/data.xls", sheet = "sub_over_vtm")
sub_under_gai <- read_excel("input/data.xls", sheet = "sub_under_gai")
sub_under_vtm <- read_excel("input/data.xls", sheet = "sub_under_vtm")

datasets = list(main_gai, main_vtm, sub_over_gai, sub_over_vtm, sub_under_gai, sub_under_vtm)
```

# Intraclass correlation  

Important references: Psychol Methods. 1996;1:30–46.   

Intraclass correlation (ICC): $ICC=\frac{Var_{Between-group}}{Var_{Total}}=\frac{\tau^2}{\tau^2 + \sigma^2}$ ($\tau$: between-group variance, $\sigma$: within-group wariance)  

Intra-rater reliability reflects that measured by a rater across multiple observations while inter-rater reliability reflects the variation of measurements between multiple raters. We are interested in the absolute agreement(GAI and VTM, respectively). In this study, there were three levels (patient N = 50, rater N = 4, and time N =2).        

# Intra-rater reliability  

*Two-way mixed-effects model (ICC[3,4])*    

$Y_{i,j,t} = \alpha + \beta t + u_i + v_ij + \epsilon_{ijt}$ ($i$: patient, $j$: time, $t$: rater, $\alpha$: intercept)  
where $u_i \sim N(0,\tau^2)$, $v_ij \sim N(0,\delta^2)$ and $\epsilon_{ijt} \sim N(0,\sigma^2)$  
$ICC=\frac{\tau^2}{\tau^2 + \frac{\delta^2+\sigma^2}{4}}$

```{r}
intra_summary_results <- data.frame(matrix(ncol = 3, nrow = length(datasets)))
colnames = c("Point estimate", "lower confidence interval", "upper confidence interval")
rownames = c("Main analysis GAI", "Main analysis VTM", "Subgroup of more than or equal to 80 years GAI", "Subgroup of more than or equal to 80 years VTM", "Subgroup of under 80 years GAI", "Subgroup of under 80 years VTM")
intra_summary_results <- setNames(intra_summary_results, colnames)
row.names(intra_summary_results) <- rownames

intra_boot_func <-  function(data, index){
  fit_intra <- lmer(angle ~ 1 + rater + (1|time/id),
                    data = data[index,],
                    control=lmerControl(check.nobs.vs.nlev = "ignore",
                                        check.nobs.vs.rankZ = "ignore",
                                        check.nobs.vs.nRE="ignore"))
  sigma_id_time_intra <- attr(VarCorr(fit_intra)[["id:time"]],"stddev")
  sigma_time_intra <- attr(VarCorr(fit_intra)[["time"]],"stddev")
  sigma_error_intra <- attr(VarCorr(fit_intra),"sc")
  icc_intra <- sigma_id_time_intra^2/(sigma_error_intra^2/4+sigma_time_intra^2/4+sigma_id_time_intra^2)
  return(icc_intra)
}

set.seed(1234)
r = 1000
for (i in 1:length(datasets)) {
  res_intra <- boot(intra_boot_func,
                  data = datasets[[i]],
                  R = r)
  intra_summary_results[i,c(1,2,3)] <- quantile(res_intra[["t"]], c(0.5, 0.025, 0.975))
}
```

# Inter-rater reliability  

*Two-way mixed-effects model (ICC[3,2])*      

$Y_{i,j,t} = \alpha + \beta k + u_i + v_{ij} + \epsilon_{ijk}$ ($i$: patient, $j$: rater, $k$: time, $\alpha$: intercept) 
where $u_i \sim N(0,\tau^2)$, $v_{ij} \sim N(0,\delta^2)$ and $\epsilon_{i,j,k} \sim N(0,\sigma^2)$  
$ICC=\frac{\tau^2}{\tau^2 + \delta^2/2 + \sigma^2/2}$

```{r}
inter_boot_func <-  function(data, index){
  fit_inter <- lmer(angle ~ 1 + time + (1|rater/id),
                    data = data[index,],
                    control=lmerControl(check.nobs.vs.nlev = "ignore",
                                        check.nobs.vs.rankZ = "ignore",
                                        check.nobs.vs.nRE="ignore"))
  sigma_id_rater <- attr(VarCorr(fit_inter)[["id:rater"]],"stddev")
  sigma_rater <- attr(VarCorr(fit_inter)[["rater"]],"stddev")
  sigma_error_inter <- attr(VarCorr(fit_inter),"sc")
  icc_inter <- sigma_id_rater^2/(sigma_error_inter^2/2+sigma_id_rater^2+sigma_rater^2/2)
  return(icc_inter)
}

inter_summary_results <- data.frame(matrix(ncol = 3, nrow = length(datasets)))
inter_summary_results <- setNames(inter_summary_results, colnames)
row.names(inter_summary_results) <- rownames

set.seed(1234)
for (i in 1:length(datasets)) {
  res_inter <- boot(inter_boot_func,
                  data = datasets[[i]],
                  R = r)
  inter_summary_results[i,c(1,2,3)] <- quantile(res_inter[["t"]], c(0.5, 0.025, 0.975))
}
```

# Difference of intra-rater reliability between GAI and VTM  

```{r}

diff <- c()

for(i in seq_len(r)){
  set.seed(i)
  data1 <- main_gai[sample(dim(main_gai)[1],dim(main_gai)[1],replace=TRUE),] 
  fit_intra1 <- lmer(angle ~ 1 + (1|time/id),
                    data = data1,
                    control=lmerControl(check.nobs.vs.nlev = "ignore",
                                        check.nobs.vs.rankZ = "ignore",
                                        check.nobs.vs.nRE="ignore"))
  sigma_id_time_intra1 <- attr(VarCorr(fit_intra1)[["id:time"]],"stddev")
  sigma_time_intra1 <- attr(VarCorr(fit_intra1)[["time"]],"stddev")
  sigma_error_intra1 <- attr(VarCorr(fit_intra1),"sc")
  icc_intra1 <- sigma_id_time_intra1^2/(sigma_error_intra1^2/4+sigma_time_intra1^2/4+sigma_id_time_intra1^2)
  
  data2 <- main_vtm[sample(dim(main_vtm)[1],dim(main_vtm)[1],replace=TRUE),] 
  fit_intra2 <- lmer(angle ~ 1 + (1|time/id),
                    data = data2,
                    control=lmerControl(check.nobs.vs.nlev = "ignore",
                                        check.nobs.vs.rankZ = "ignore",
                                        check.nobs.vs.nRE="ignore"))
  sigma_id_time_intra2 <- attr(VarCorr(fit_intra2)[["id:time"]],"stddev")
  sigma_time_intra2 <- attr(VarCorr(fit_intra2)[["time"]],"stddev")
  sigma_error_intra2 <- attr(VarCorr(fit_intra2),"sc")
  icc_intra2 <- sigma_id_time_intra2^2/(sigma_error_intra2^2/4+sigma_time_intra2^2/4+sigma_id_time_intra2^2)
  
  diff <- c(diff, icc_intra1-icc_intra2)
}

diff_intra_main <- quantile(diff,c(0.5, 0.025, 0.975))
```

# Difference of inter-rater reliability between GAI and VTM  

```{r}

diff <- c()

for(i in seq_len(r)){
  set.seed(i)
  data1 <- main_gai[sample(dim(main_gai)[1],dim(main_gai)[1],replace=TRUE),] 
  fit_inter1 <- lmer(angle ~ 1 + (1|rater/id),
                    data = data1,
                    control=lmerControl(check.nobs.vs.nlev = "ignore",
                                        check.nobs.vs.rankZ = "ignore",
                                        check.nobs.vs.nRE="ignore"))
  sigma_id_rater_inter1 <- attr(VarCorr(fit_inter1)[["id:rater"]],"stddev")
  sigma_rater_inter1 <- attr(VarCorr(fit_inter1)[["rater"]],"stddev")
  sigma_error_inter1 <- attr(VarCorr(fit_inter1),"sc")
  icc_inter1 <- sigma_id_rater_inter1^2/(sigma_error_inter1^2/2+sigma_rater_inter1^2/2+sigma_id_rater_inter1^2)
  
  data2 <- main_vtm[sample(dim(main_vtm)[1],dim(main_vtm)[1],replace=TRUE),] 
  fit_inter2 <- lmer(angle ~ 1 + (1|rater/id),
                    data = data2,
                    control=lmerControl(check.nobs.vs.nlev = "ignore",
                                        check.nobs.vs.rankZ = "ignore",
                                        check.nobs.vs.nRE="ignore"))
  sigma_id_rater_inter2 <- attr(VarCorr(fit_inter2)[["id:rater"]],"stddev")
  sigma_rater_inter2 <- attr(VarCorr(fit_inter2)[["rater"]],"stddev")
  sigma_error_inter2 <- attr(VarCorr(fit_inter2),"sc")
  icc_inter2 <- sigma_id_rater_inter2^2/(sigma_error_inter2^2/2+sigma_rater_inter2^2/2+sigma_id_rater_inter2^2)
  
  diff <- c(diff, icc_inter1-icc_inter2)
}

diff_inter_main <- quantile(diff,c(0.5, 0.025, 0.975))
```

# Extra: inter-rater reliabliity of three variance components  

$Y_{i,j,t} = \alpha + u_i + v_{ij} + z_{ijk} + \epsilon_{ijk}$ ($i$: patient, $j$: rater, $k$: time, $\alpha$: intercept) 
where $u_i \sim N(0,\tau^2)$, $v_{ij} \sim N(0,\delta^2)$, $z_{ijk} \sim N(0,\omega^2)$ and $\epsilon_{i,j,k} \sim N(0,\sigma^2)$  
$ICC=\frac{\tau^2}{\tau^2 + \delta^2 + \omega^2 + \sigma^2}$

However, this model is meaning less considering the number of levels and sample size.    

```
inter_boot_func2 <-  function(data, index){
  fit_inter <- lmer(angle ~ 1 + (1|time/rater/id),
                    data = data[index,],
                    control=lmerControl(check.nobs.vs.nlev = "ignore",
                                        check.nobs.vs.rankZ = "ignore",
                                        check.nobs.vs.nRE="ignore"))
  sigma_id_rater_time <- attr(VarCorr(fit_inter)[["id:(rater:time)"]],"stddev")
  sigma_rater_time <- attr(VarCorr(fit_inter)[["rater:time"]],"stddev")
  sigma_time <- attr(VarCorr(fit_inter)[["time"]],"stddev")
  sigma_error_inter <- attr(VarCorr(fit_inter),"sc")
  icc_inter <- (sigma_id_rater_time^2)/(sigma_error_inter^2 + sigma_id_rater_time^2 + sigma_rater_time^2 + sigma_time^2)
  return(icc_inter)
}

inter_summary_results2 <- data.frame(matrix(ncol = 3, nrow = length(datasets)))
inter_summary_results2 <- setNames(inter_summary_results2, colnames)
row.names(inter_summary_results2) <- rownames

set.seed(1234)
for (i in 1:length(datasets)) {
  res_inter2 <- boot(inter_boot_func2,
                  data = datasets[[i]],
                  R = r)
  inter_summary_results[2i,c(1,2,3)] <- quantile(res_inter2[["t"]], c(0.5, 0.025, 0.975))
}
```

# Results  

## Intra-rater reliablilty  

```{r}
kable(intra_summary_results)
```

## Inter-rater reliablilty  

```{r}
kable(inter_summary_results)
```

## Difference in intra-rater reliability  

```{r}
kable(diff_intra_main)
```

## Difference in inter-rater reliability  

```{r}
kable(diff_inter_main)
```

