---
title: "Yamamoto Study"
author: "Akihiro Shiroshita"
date: "`r Sys.time()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	fig.height = 4,
	fig.pos = "t",
	message = FALSE,
	warning = FALSE,
	dpi = 350,
	out.extra = ""
)
packages = c("data.table",
             "lava",
             "lme4",
             "nlme",
             "readxl",
             "irr",
             "tidyverse",
             "psych",
             "lubridate",
             "ggplot2",
             "RColorBrewer",
             "ggplotgui",
             "boot")
package.check <- lapply(packages, FUN = function(x){
  if (!require(x, character.only = TRUE)){
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

gai_data <- read_excel("input/gai.xls")
```

# Intraclass correlation  

Important references: Psychol Methods. 1996;1:30–46.   

Intraclass correlation (ICC): $ICC=\frac{Var_{Between-group}}{Var_{Total}}=\frac{\tau^2}{\tau^2 + \sigma^2}$ ($\tau$: between-group variance, $\sigma$: within-group wariance)  

Intra-rater reliability reflects that measured by a rater across multiple observations while inter-rater reliability reflects the variation of measurements between multiple raters. We are **not** interested in the absolute agreement, but the consistency of the methodology (GAI and VTM, respectively). In this study, there were three levels (patient N = 50, rater N = 4, and time N = 2).        

# Intra-rater reliability  
This analysis aimed to generalize the intra-rater reliability to general raters. 
Thus, we are interested in specific raters, but are not interested in time.  
Three-way mixed-effects model.    

$Y_{i,j,t} = \alpha + \beta_t t + u_i + v_ij + \epsilon_{i,j,t}$ ($t$: rater, $i$: patient, $j$: time, $\alpha$: intercept)  
where $u_i \sim N(0,\tau^2)$, $v_{ij} \sim N(0,\delta^2)$ and $\epsilon_{i,j,t} \sim N(0,\sigma^2)$  
$ICC=\frac{\tau^2}{\tau^2 + \frac{\sigma^2 + \delta^2}{4}}$

```{r}
fit_intra <- lmer(angle ~ 1 + rater + (1|id/time), data = gai_data)

sigma_id_intra <- attr(VarCorr(fit_intra)[["id"]],"stddev")
sigma_id_time_intra <- attr(VarCorr(fit_intra)[["time:id"]],"stddev")
sigma_rater_intra <- attr(VarCorr(fit_intra)[["rater"]],"stddev")
sigma_error_intra <- attr(VarCorr(fit_intra),"sc")

icc_intra <- sigma_id_intra^2/((sigma_error_intra^2+sigma_id_time_intra^2+sigma_id_time_intra^2)/4+sigma_id_intra^2)
icc_intra

intra_boot_func <-  function(data, index){
  fit_intra <- lmer(angle ~ 1 + rater + (1|id/time), data = data[index,])
  sigma_id_intra <- attr(VarCorr(fit_intra)[["id"]],"stddev")
  sigma_id_time_intra <- attr(VarCorr(fit_intra)[["time:id"]],"stddev")
  sigma_rater_intra <- attr(VarCorr(fit_intra)[["rater"]],"stddev")
  sigma_error_intra <- attr(VarCorr(fit_intra),"sc")
  icc_intra <- sigma_id_intra^2/((sigma_error_intra^2+sigma_id_time_intra^2)/4+sigma_id_intra^2)
  return(icc_intra)
}

set.seed(1234)
res_intra <- boot(intra_boot_func,
                  data = gai_data,
                  R = 1000)
res_intra[["t0"]]
boot.ci(res_intra, type = "perc")
```

# Inter-rater reliability  

This analysis aimed to generalize the inter-rater reliability to general raters. 
Thus, we are not interested in specific raters. Actually, the raters were *not* randomly selected from the source population.    
Three-way random-effects model.    

The variance components splits the variance into three pieces (patient, rater, time).
$Y_{i,j,t} = \alpha + u_i + v_j + z_k + \epsilon_{i,j,k}$ ($i$: patient, $j$: time, $k$: rater, $\alpha$: intercept) 
where $u_i \sim N(0,\tau^2)$, $v_j \sim N(0,\delta^2)$, $z_k \sim N(0,\omega^2)$ and $\epsilon_{i,j,k} \sim N(0,\sigma^2)$  
$ICC=\frac{\tau^2}{\tau^2 + \delta^2 + \sigma^2 + \omega^2}$

```{r}
fit_inter <- lmer(angle ~ 1 + (1|time) + (1|id) + (1|rater), data = gai_data)

sigma_id_inter <- attr(VarCorr(fit_inter)[["id"]],"stddev")
sigma_time_inter <- attr(VarCorr(fit_inter)[["time"]],"stddev")
sigma_rater_inter <- attr(VarCorr(fit_inter)[["rater"]],"stddev")
sigma_error_inter <- attr(VarCorr(fit_inter),"sc")

icc_inter <- sigma_id_inter^2/(sigma_time_inter+sigma_rater_inter+sigma_id_inter^2)
icc_inter

inter_boot_func <-  function(data, index){
  fit_inter <- lmer(angle ~ 1 + (1|time) + (1|id) + (1|rater), data = data[index,])
  sigma_id_inter <- attr(VarCorr(fit_inter)[["id"]],"stddev")
  sigma_time_inter <- attr(VarCorr(fit_inter)[["time"]],"stddev")
  sigma_rater_inter <- attr(VarCorr(fit_inter)[["rater"]],"stddev")
  sigma_error_inter <- attr(VarCorr(fit_inter),"sc")
  icc_inter <- sigma_id_inter^2/(sigma_error_inter^2+sigma_time_inter+sigma_rater_inter+sigma_id_inter^2)
  return(icc_inter)
}

set.seed(1234)
res_inter <- boot(inter_boot_func,
                  data = gai_data,
                  R = 1000)
res_inter[["t0"]]
boot.ci(res_inter, type = "perc")
```
